Illuminating a Dark Lens : A Type Ia Supernova Magnified by Galaxy Cluster Abell 2744

### 2015-02-18 snTomas.tex
added light curve fitting figure and table  showing the mlcs2k2 and salt2 fits

\author{
 Steven~A.~Rodney\altaffilmark{1},
  \etal
}
\altaffiltext{1}{\JHU}
\altaffiltext{2}{Hubble fellow}

@@

Lensed SNe have been discovered
\citep{Goobar:2009,Riehm:2011,Patel:2014,Nordin:2014}

@@

As a check on the spectral classification of SN \tomas\
(Section~\ref{sec:SNID}), we independently classified the SN using a
Bayesian photometric classifier.  This classification code uses the
{\tt sncosmo} software
package\footnote{\url{http://sncosmo.github.io/}} to simulate SN light
curves at a range of redshifts and evaluate the classification
probability using traditional Bayesian model selection \citep[as
in][]{Jones:2013,Rodney:2014,Graur:2014,Rodney:2015}.  In this
analysis we represent normal Type Ia SNe with the SALT2
model \citep{Guy:2010}, and \CCSNe\ with 42 discrete templates (26
Type II and 16 Type Ib/c) drawn from the template library of the
SuperNova Analysis software package \citep[SNANA,][]{Kessler:2009a}.
Likelihoods are defined by comparing the observed fluxes to model
predictions in all passbands where the model is defined.  In practice,
this means we exclude the F435W and F606W bands, which are too blue
for our models at $z>1$.

The \CCSN\ models have free parameters for date of peak brightness
($t_{\rm pk}$), amplitude, and redshift ($z$). Due to the expected
impact of gravitational lensing magnification, we do not include any
prior on the intrinsic luminosity for any SN sub-class.  We also do
not assign a prior for the SN redshift. This allows our photometric
analysis to provide an independent check on the host galaxy photo-$z$
and spectroscopic redshift (Sections~\ref{sec:HostGalaxy}
and \ref{sec:SpecFit}).

@@

The final photometric classification probability for SN \tomas\ is
$p({\rm Ia}|{\bf \rm D})=1.0$, with the classification
probability from all \CCSN\ sub-classes totaling less than $10^{-100}$.  The
marginal posterior distribution in redshift for the Type Ia model is
sharply peaked at $z=1.36\pm0.02$, which is $<2.5\sigma$ from
the spectroscopic redshift of $z=1.31\pm0.01$ derived in
Section~\ref{sec:SpecRedshift}.  The time of peak brightness is also
tightly constrained at $t_{\rm pk}=56816.0\pm0.2$, which means the
spectroscopic observations were collected within 2 rest-frame days of
the epoch of peak brightness, consistent with our spectroscopic
analysis.

@@

With the spectroscopic type and redshift securely defined as a normal
Type Ia SN at $z=1.3$, we now turn to fitting the light curve with
Type Ia templates to measure the distance modulus.  Here we use two
independent light curve fitters: first the SALT2 model described
above, and then the MLCS2k2 model from \citep{Jha:2007}.

\subsection{SALT2 Fit}
\label{sec:SALT2Fit}

in \citep{Patel:2014}:   $\alpha=1.367\pm0.086$,
$\beta=3.179\pm0.101$, $M=-19.12\pm0.03$.

For the SALT2 Type Ia model one can define a shape- and color-corrected apparent
B-band magnitude using 
$m_{\rm B}^\prime = m_{\rm B} + \alpha x_1 - \beta c$.

For the SALT2 model we fix the model parameters to the values
derived in the \citet{Rest:2014} analysis of the Pan-STARRS1 SN sample:
$\alpha=0.147$, $\beta=3.13$, $M_B=-19.34$.

Comparing salt2 color to Av \citep{Kessler:2009b}:
$c = (0.464 ± 0.021)AV − (0.121 ± 0.014)$

from \citet{Patel:2014}:
``Finally, to ensure a consistent zeropoint between the SALT2 and
MLCS2k2 distance moduli, we applied both methods to a sample of SDSS
Supernova Survey SNe Ia from Holtzman et al. (2008) and Kessler et
al. (2009a). We added a zeropoint correction of 0.20 mag to the
distance moduli reported by SNANA in the MLCS2k2 fit (dmMLCS2k2), to
put them on the same scale as the SALT2 distances (dmSALT2).''

\subsection{MLCS2k2 Fit}
\label{sec:MLCS2k2Fit}

As in \citep{Patel:2014}.

$R_v$ fixed to 1.9.

for Rv = 1.9, the calibrated distance is: $\mu = \mu_{\rm mlcs}-0.2006$.

\begin{figure}
\begin{center}
\includegraphics[width=0.49\textwidth]{FIG/lcfit_mlcs2k2_ABmag}
\caption{  Light curve fit using MLCS2k2. 
\label{fig:mlcsfit} }
\end{center}
\end{figure}

\section{Lensing Magnification}
\label{sec:LensingMagnification}

\subsection{Direct Magnification Measurement}
\label{sec:DirectMagnificationMeasurement}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/magnification_measurement}
\caption{ Deriving the magnification from comparison to field SN
\label{fig:magmeasure} }
\end{center}
\end{figure}

>>>>

\author{
 Steven~A.~Rodney\altaffilmark{1,2},
  \etal
}
\altaffiltext{1}{\JHU}
\altaffiltext{2}{Hubble fellow}

@@

Lensed SNe have been discovered
\citep{Goobar:2009,Riehm:2011}

\citet[][hereafter P14]{Patel:2014} and \citet{Nordin:2014} presented
independent analyses of three lensed SNe, of which at least 2 are securely
classified as Type Ia SNe.  All of these three objects were found in
the Cluster Lensing and Supernova survey with Hubble (CLASH,
PI:Postman, HST Program ID 12068).

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/snTomas_lightcurve_classification}
\caption{ 
Maximum likelihood model for each SN sub-class, derived from Bayesian
model selection using the photometric data alone. Grey points show the
observed SN \tomas\ photometry with error bars, though these are typically
smaller than the size of the marker.  The Type Ia model
(orange solid line) is drawn from the SALT2 template at $z=1.35$. The
best match from all Type Ib/Ic models is based on the Type Ic SN
SDSS-14475 at $z=0.695$ (green dashed line). For the Type II class, the
best match is from the Type II-L SN 2007pg at $z=1.8$ (blue dash-dot line). The Type
Ia model is by far the best match, and the only one that is consistent
with both the photo-$z$ of the probable host galaxy and the
spectroscopic redshift from the SN spectrum.
\label{fig:photoclass} }
\end{center}
\end{figure}

As a check on the spectral classification of SN \tomas\
(Section~\ref{sec:SNID}), we independently classified the SN using a
Bayesian photometric classifier.  We use the
{\tt sncosmo} software
package\footnote{\url{http://sncosmo.github.io/}} to simulate SN light
curves from $z=0.3$ to 2.3 and evaluate the classification
probability using traditional Bayesian model selection \citep[as
in][]{Jones:2013,Rodney:2014,Graur:2014,Rodney:2015}.  In this
analysis we represent normal Type Ia SNe with the SALT2
model \citep{Guy:2010}, and \CCSNe\ with 42 discrete templates (26
Type II and 16 Type Ib/c) drawn from the template library of the
SuperNova Analysis software
package \citep[SNANA,][]{Kessler:2009a}.\footnote{Throughout this work
we use SNANA v10\_35g.}
Likelihoods are defined by comparing the observed fluxes to model
predictions in all passbands where the model is defined.  In practice,
this means we exclude the F435W and F606W bands, which are too blue
for our models at $z>0.85$.

The \CCSN\ models have free parameters for date of peak brightness
($t_{\rm pk}$), amplitude, and redshift ($z$). Due to the expected
impact of gravitational lensing magnification, we do not include any
prior on the intrinsic luminosity for any SN sub-class.  We also do
not assign a prior for the SN redshift. This allows our photometric
analysis to provide an independent check on the host galaxy photo-$z$
and spectroscopic redshift (Sections~\ref{sec:DiscoveryAndFollowup}
and \ref{sec:SpecRedshift}).

@@

The final photometric classification probability for SN \tomas\ is
$p({\rm Ia}|{\rm {\bf D}})=1.0$, with the classification probability
from all \CCSN\ sub-classes totaling less than $10^{-32}$.  The
best-fit models for each sub-class are shown in
Figure~\ref{fig:photoclass}, which demonstrates that the \CCSN\ models
cannot adequately match the observed photometry. In particular, only
the Type Ia model can simultaneously provide an acceptable fit to the
well-sampled rising light curve in F814W and the F814W-F160W
color near peak.

The marginal posterior distribution in redshift for the Type Ia model
is sharply peaked at $z=1.35\pm0.02$, which is fully consistent with
the photo-$z$ of the presumed host galaxy ($z=1.5\pm0.2$), and
$<2\sigma$ from the spectroscopic redshift of $z=1.31\pm0.01$ derived
in Section~\ref{sec:SpecRedshift}.  The time of peak brightness is
also tightly constrained at $t_{\rm pk}=56816.3\pm0.3$, which means
the spectroscopic observations were collected within 2 rest-frame days
of the epoch of peak brightness, consistent with our spectroscopic
analysis.

@@

With the spectroscopic type and redshift securely defined as a normal
Type Ia SN at $z=1.3$, we now turn to fitting the light curve with
Type Ia templates to measure the distance modulus.  Here we use two
independent light curve fitters: the SALT2 model described
above and the MLCS2k2 model \citep{Jha:2007}.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{FIG/snTomas_lightcurve_fit_fluxAB25}
\caption{ Type Ia light curve fits to SN \tomas\ using the MLCS2k2
(top row) and SALT2 (bottom row) fitters. The model redshifts are set to
$z=1.33\pm0.02$, encompassing both redshift values as determined from
spectroscopic and photometric constraints.  Solid lines denote the
best-fit model and shaded lines show the range allowed by 1-$\sigma$
uncertainties on the model parameters. 
Observed fluxes are shown as diamonds, scaled to an AB magnitude
zero point of 25. Error bars are plotted, but most are commensurate
with the size of the points. The left-most panel includes observations
in the F435W and F606W filters, although these were not used for the
fit, as they are bluer than the minimum wavelength for the model. The
lower axis marks time in observer frame days, while the top axis shows
the time in the rest frame relative to the epoch of peak brightness.
\label{fig:LightCurveFits} }
\end{center}
\end{figure*}

With both fitters we find light curve shape and color parameters that
are fully consistent with a normal Type Ia SN, regardless of whether
we adopt the spectroscopic redshift $z=1.31$ from
Section~\ref{sec:SpecRedshift} or the photometric redshift $z=1.35$
from Section~\ref{sec:PhotometricClassification}.  For SALT2, with the
redshift range set to $1.33\pm0.02$, we find a light curve shape
parameter of $x_1=0.164\pm0.199$ and a color parameter of
$c=-0.115\pm0.025$, yielding a $\chi^2$ value of 48.4 for 36 degrees
of freedom, $\nu$.  With the MLCS2k2 fitter the best-fit shape
parameter is $\Delta=-0.033\pm0.083$ and the color term is
$A_{\rm V}=0.014\pm0.037$, giving $\chi^2/\nu=23.9/36$.

\subsection{Distance Modulus}
\label{sec:DistanceModulus}

As in \citet{Patel:2014}, we derive a distance modulus 
from the SALT2 fit using\footnote{This avoids confusion by reserving the symbol $\mu$
to refer to the lensing magnification. The 'dm' is a standard distance
modulus, defined as ${\rm dm}=5\log_{10}d_L+25$, where $d_L$ is the
luminosity distance in Mpc.}

\begin{equation} \label{eqn:dmSALT2}
 {\rm dm}_{\rm SALT2} = m_B^* - M + \alpha(s-1) - \beta C.
\end{equation}

\noindent  Here the parameters for light curve shape $s$ and color $C$ 
correspond to the SiFTO light curve fitter \citep{Conley:2008}, so we
first use the formulae from \citet{Guy:2010} to convert from SALT2
($x_1$ and $c$) into the equivalent SiFTO parameters.  We also add an
offset of 0.27 mag to the value of $m_B^*$ returned by SNANA, in order
to match the arbitrary normalization of the SALT2 fitter used
by \citet{Guy:2010} and \citet{Sullivan:2011}.  This conversion allows
us to adopt values for the constants $M$, $\alpha$, and $\beta$
from \citet{Sullivan:2011}, which have been calibrated using 472 SNe
from the SNLS3 sample \citep{Conley:2011}: $M=-19.12\pm0.03$,
$\alpha=1.367\pm0.086$, and $\beta=3.179\pm0.101$.

\input{TABLE/lcfits}

The SNANA version of the MLCS2k2 fitter returns a value for the
distance modulus (dm$_{\rm MLCS2k2}$) that has an arbitrary zero point
offset relative to the SALT2 distances (dm$_{\rm SALT2}$).  To put the
two distances onto the same reference frame we add a zeropoint
correction of 0.20 mag to the MLCS2k2 distances as
in \citet{Patel:2014}.  This correction was derived by applying both
fitters to a sample of Type Ia SNe from the SDSS
survey \citep{Holtzman:2008,Kessler:2009b}, with the extinction law
$R_V$ fixed at 1.9.

Final values for the SN distance modulus are shown in
Table~\ref{tab:DistanceAndMagnification}, both for the spectroscopic
redshift ($z=1.31$) and the photometric redshift ($z=1.35$).  The two
light curve fitters are consistent within the uncertainties.

\subsection{Magnification}
\label{sec:Magnification}

To measure the lensing magnification, we would like to avoid
introducing systematic uncertainties inherent to any assumed
cosmological model \citep[e.g.][]{Nordin:2014}.  To that end, we
follow \citet{Patel:2014} and define the magnification by comparing
the measured distance modulus of SN \tomas\ against an average
distance modulus derived from a ``control sample'' of unlensed Type Ia
SN at similar redshift.  This allows us to make only the minimal
assumption that the redshift-distance relationship for Type Ia SN is
smooth and approximately linear over a small redshift span, which
should be true for any plausible cosmological model.

The unlensed sample comprises 18 spectroscopically confirmed Type Ia
SNe in the range $1.14\leq z \leq1.42$ from the GOODS and SCP
surveys,\footnote{GOODS: Great Observatories Origins Deep Survey,
PI:Giavalisco, HST-PID:; SCP: Supernova Cosmology Project,
PI:Perlmutter} which used the HST Advanced Camera for
Surveys \citep{Riess:2006,Suzuki:2012}.  Using the SALT2 and MLCS2k2
fitters as described above, we get distance modulus measures for every
object in this control sample.  We then fit a linear relationship for
distance modulus vs. redshift, and derive a prediction for the
distance modulus of a normal Type Ia SN at the redshift of SN \tomas
(Figure~\ref{fig:MagnificationMeasurement}).
This predicted value is given in
Table~\ref{tab:DistanceAndMagnification} under the ``Control'' column.
The difference between the observed distance of SN \tomas\ and this
control sample value is attributed to the magnification from
gravitational lensing:

\begin{equation} \label{eqn:mu}
{\rm dm}_{\rm control} - {\rm dm}_{\rm \tomas} = 2.5 \log_{10}\mu.
\end{equation}

\noindent This inferred magnification is reported in the final column of
Table~\ref{tab:DistanceAndMagnification}.

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{FIG/magnification_measurement}
\caption{ Measurement of the lensing magnification from comparison of
the \tomas\ distance modulus to a sample of unlensed field SN.
The distance modulus for each SN is derived from light curve fits
using the SALT2 fitter (top panel) and the MLCS2k2 fitter (right).
\label{fig:MagnificationMeasurement} }
\end{center}
\end{figure}
